---
title: "IEAP - Rstudio Series 02"
author: "Zoe Sauge, Elouan DEFFIN, Tuba"
date: last-modified
date-format: "DD MMMM YYYY, HH:mm:ss"
format:
  typst:
    papersize: a4
    margin:
      x: 1.5cm
      y: 1.2cm
    fontsize: 9.5pt
---

## 2. Statistical tests: comparison of medians (or means?)

### 2.1 Make clear what is a statistical test, and when to use it.

**What is a median? A mean?** The mean is the sum of all values divided by their number, and the median is the middle value once the data are sorted. *Reference: Field, A. (2013). Discovering Statistics Using IBM SPSS Statistics, Chapter 2.*

**What is a variance? A standard deviation?** The variance is the average squared deviation from the mean, and the standard deviation is its square root, expressed in the same unit as the data. *Reference: Field, A. (2013). Discovering Statistics Using IBM SPSS Statistics, Chapter 2.*

**What is a normal distribution?** A normal distribution is a symmetric, bell-shaped distribution in which most values cluster around the mean. *Reference: STHDA - Normality Test in R*

**What is a statistical test?** A statistical test uses sample data to decide whether there is enough evidence to reject a null hypothesis, by computing a p-value. *Reference: STHDA - Statistical Tests and Assumptions*

**When to use a statistical test?** A statistical test is used to know whether a result observed on a sample is likely real or due to random sampling variation. *Reference: STHDA - Statistical Tests and Assumptions*

**What is a parametric test? A non-parametric test?** A parametric test assumes a specific distribution of the data (usually normal), whereas a non-parametric test makes no such assumption and typically works on ranks. *Reference: STHDA - Non-parametric Tests*

**What are the assumptions of parametric tests?** Parametric tests generally assume normally distributed data, independent observations and, for group comparisons, equal variances. *Reference: Field, A. (2013). Discovering Statistics Using IBM SPSS Statistics, Chapter 5.*

**What are the assumptions of non-parametric tests?** Non-parametric tests mainly assume independent observations and data that can be ranked. *Reference: STHDA - Non-parametric Tests*

**What is a p-value?** The p-value is the probability of obtaining a result at least as extreme as the one observed if the null hypothesis is true. *Reference: STHDA - Statistical Tests and Assumptions*

**What is the risk of error when using a statistical test?** A Type I error is a false positive (risk usually set at α = 0.05), and a Type II error is a false negative, often due to low statistical power. *Reference: Field, A. (2013). Discovering Statistics Using IBM SPSS Statistics, Chapter 2.*

**What is the difference between a paired and an unpaired test?** A paired test compares two measurements from the same subjects, whereas an unpaired test compares two independent groups. *Reference: STHDA - t-test Formula*

## 2.2 Effect of treatment over time

### 2.2.1 The data 

```{r}
library(tidyverse)
```


```{r}
prepost <- read_csv("data/PrePost.csv")

glimpse(prepost)
```
```{r}
summary(prepost)
```

```{r}
count(prepost, time)
```

The dataset contains 16 observations: 8 measurements before rehabilitation and 8 measurements after rehabilitation. The variable time identifies the measurement period (Before or After), while perf represents the measured performance variable.

| Variable | Description | Unit | Possible values | Expected change |
|---------------|---------------|---------------|---------------|---------------|
| perf | Cardiac rehabilitation performance score; the exact test/measure is not specified in the assignment materials | Not specified (likely arbitrary or composite score) | 125–148 observed in this dataset | Expected to improve after rehabilitation |
| time | Measurement period | Categorical | Before / After | Not applicable |

### 2.2.2 The analysis

```{r}
ggplot(prepost, aes(x = time, y = perf)) +
  geom_boxplot() +
  geom_jitter(width = 0.1) +
  labs(
    title = "Performance Before and After Rehabilitation",
    x = "Time",
    y = "Performance"
  )
```

```{r}
prepost
```

```{r}
before <- prepost$perf[prepost$time == "Before"]
after <- prepost$perf[prepost$time == "After"]

difference <- after - before

difference
```

```{r}
shapiro.test(difference)
```

A paired t-test was used because the same participants were measured before and after rehabilitation. The distribution of the paired differences did not show evidence of departure from normality according to the Shapiro-Wilk test (p = 0.283).

```{r}
t.test(after, before, paired = TRUE)
```

The paired t-test showed a statistically significant increase in performance after rehabilitation (t(7) = 5.40, p = 0.001). Performance increased by an average of 2.5 units, with a 95% confidence interval from 1.41 to 3.59 units. These results suggest that performance improved after the rehabilitation program.

